---
title: "hip_replacement_operations"
author: "XI CHEN"
date: "2024-10-30"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim

Plot 'EQ-5D Index' scores (a combination fo five key criteria concerning patients' self-reported general health) pre and post a hip replacement operation for different age groups.

## Load packages

```{r load-packages}
library(tidyverse)
```

## Read in data

The data is in the file "Hip Replacement CCG 1819.csv", and it contains patient reported outcomes for hip replacement procedures, form April 2018 to March 2019.
It was downloaded from https://digital.nhs.uk/data-and-information/publications/statistical/patient-reported-outcome-measures-proms/for-hip-and-knee-replacement-procedures-april-2018-to-march-2019
We also have the data dictionary for this dataset in "proms_data_diciionary.pdf".

```{r}
hip_data <- read.csv("Data/Hip Replacement CCG 1819.csv")
```

## Prepare the data

```{r inspect-data}
glimpse(hip_data)
```

## Select age and quality of life score pre and post operation

```{r select-age-eq5d}
age_EQ5D <- hip_data %>% 
  select('Age.Band','Pre.Op.Q.EQ5D.Index','Post.Op.Q.EQ5D.Index') %>% 
  rename(Age = 'Age.Band',
         EQ5D_Pre = 'Pre.Op.Q.EQ5D.Index',
         EQ5D_Post = 'Post.Op.Q.EQ5D.Index'
         )

head(age_EQ5D)
```
## Remove missing values

```{r missing-values}
age_EQ5D$Age %>% unique()
age_EQ5D$Age %>% table()
age_EQ5D %>% summary()

age_EQ5D_noNA <- age_EQ5D %>% 
  drop_na() %>% 
  filter(Age !='*')

table(age_EQ5D_noNA$Age)
summary(age_EQ5D_noNA)
```

## Check that data is tidy

```{r tidy-data}
head(age_EQ5D_noNA)

tidy_age_EQ5D_noNA <- age_EQ5D_noNA %>% 
  pivot_longer(c(EQ5D_Pre,EQ5D_Post),
               names_to = 'Time',
               names_prefix = 'EQ5D_',
               values_to = 'EQ5D'
               )
head(tidy_age_EQ5D_noNA)
```

## Plot quality of life pre and operation for each age group

```{r plot-eq5d}
tidy_age_EQ5D_noNA$Time <- factor(tidy_age_EQ5D_noNA$Time,levels=c('Pre','Post')) 

tidy_age_EQ5D_noNA %>% 
  ggplot() +
  geom_boxplot(aes(x = Time, y = EQ5D, colour = Age))
```

