---
title: "problems_cause_by_a_troke"
author: "XI CHEN"
date: "2024-10-31"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim
Calculate how many patients in this dataset have been told by a doctor that they have problems caused by a stroke.

## Load Packages
```{r load-packages}
library(tidyverse)
```

## Read in Data
The data is in the file "Hip Replacement CCG 1819.csv", and it contains patient reported outcomes for hip replacement procedures, form April 2018 to March 2019.
It was downloaded from https://digital.nhs.uk/data-and-information/publications/statistical/patient-reported-outcome-measures-proms/for-hip-and-knee-replacement-procedures-april-2018-to-march-2019
We also have the data dictionary for this dataset in "proms_data_diciionary.pdf".

```{r read-in-data}
hip_data <- read.csv("Data/Hip Replacement CCG 1819.csv")
```

## Prepare the Data

```{r inspect-data}
glimpse(hip_data)
```

## Select stroke
```{r select-stroke}
stroke <- hip_data %>% 
  select('Stroke')

head(stroke)
```
## Calculate the number of patients in this dataset who have been told by a doctor that they have problems caused by a stroke.
```{r}
table (stroke)
```
According to proms_data_dictionary.pdf, in the field of Stroke, value 1 means Yes, value 9 means Missing. From the table, we can see there are 400 patients in this dataset who have been told by a doctor that they have problems caused by a stroke. 28,520 patients are missing data.
