---
title: "table_operation_activity_levels"
author: "XI CHEN"
date: "2024-10-31"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim
Create a clean and tidy table with pre and post operation activity levels.

## Load Packages
```{r load-packages}
library(tidyverse)
```

## Read in Data
The data is in the file "Hip Replacement CCG 1819.csv", and it contains patient reported outcomes for hip replacement procedures, form April 2018 to March 2019.
It was downloaded from https://digital.nhs.uk/data-and-information/publications/statistical/patient-reported-outcome-measures-proms/for-hip-and-knee-replacement-procedures-april-2018-to-march-2019
We also have the data dictionary for this dataset in "proms_data_diciionary.pdf".
```{r read-data}
hip_data <- read.csv("Data/Hip Replacement CCG 1819.csv") 
```

## Prepare the Data
```{r inspect-data}
glimpse(hip_data)
```

## Select Pre and Post Operation Activity Levels
```{r select-pre-post-activity-levels}
activity_levels <- hip_data %>% 
  select('Pre.Op.Q.Activity', 'Post.Op.Q.Activity') %>% 
  rename(Activity_Pre = 'Pre.Op.Q.Activity',
         Activity_Post = 'Post.Op.Q.Activity'
         )
head(activity_levels)
```

## Removing Missing Data
According to proms_data_dictionary.pdf, the field of Activity, value equals to 9 means missing data.
```{r removing-data}
activity_levels_noNA <- activity_levels %>%
 drop_na() %>% 
  filter(Activity_Pre !='9', Activity_Post !='9' )
table(activity_levels_noNA)
```
## Check the Data is Tidy

```{r tidy-data}
head(activity_levels_noNA)

tidy_activity_levels_noNA <- activity_levels_noNA %>% 
  pivot_longer(c(Activity_Pre,Activity_Post),
                 names_to = 'Time',
                 names_prefix = 'Activity_',
                 values_to = 'Activity'
               )
head(tidy_activity_levels_noNA)
```
## Create a Table with Pre and Post Operation Activity Levels
```{r create-table}
table(tidy_activity_levels_noNA)
```

